Speculation in foreign exchange: noise or information

### 2015-07-11 Knitr/NewPaper.Rnw
Updated part of the NewPaper. This does not fully work.  The pdf does not come up.  It is not clear what is wrong. A pdf is created.  Need to update something on momentum trading and on the information that some traders have.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study is used to assess the views of Keynes and Friedman on speculation. Speculative extremes are ranked by intensity of sentiment and weight of speculative activity. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.  The extremes say little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity. 
\end{abstract}
\section*{Introduction}
There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of this spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that does not emerge randomly, but feeds off itself as the evolution of collective sentiment encourages trends in asset prices.

Towards the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
    The foreign exchange market is utilised because it is the largest and most liquid market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}.  The cost of trading in this market, when measurd in terms of the size of the bid-ask spread, is extremely low. An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}.

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with sentiment and positions, there is no informational content from either. As such, knowing the the strength of speculative opinion or the weight of speulative positions does not improve forecasts of future returns. This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.

\section{Literature}
\subsection*{Overview}
  Speculation is an important concept in economics, particular in the area of finance.  However, there is no clear, unambiguous theory of speculation. Speculation plays a key role in three importantant and related areas: market-microstructure, the rationale for the trading of financial securities and rational expectations.

@@

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninforrmed speculation provides a rational and an opportunity for the trading of financial securities. If all economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, the informed can take advantage of their information.  This creates a justification for acquiring information. Speulative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundametnal value is thereby essential to encourage te process of price discovery. 
%
 \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

  Speculation is also an important part of the argument for why prices should tend towards that consistent with \emph{rational expectations}.  \citet{muth1961Rational} uses commodity speculation as the means to drive prices towards the \emph{rationa-expectations-equilibrium} (RER).  In this case, the speculators are informed and tend to reduce the volatility of prices. This is a reiteration of the Friedman idea of speuclation but runs immediately into the \citet{milgrom1982information} problem of finding counterparties willing to make the other side of the transaction.

% See \citep{Lyons2006Microstructure} and \citep{Evans2002Order} for examples of the micrstrucure approach in foreign exchange.

If there are a mixture of informed and uninformed specualtors, the question becomes whether the uninformed speculative activity can be considered as just random noise?   The original microstrucutre version had moise-traders arriving in a random fashion so that the market-maker could infer information from an imbalance of orders \citet{Kyle1985Continuous}.  However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inapprpropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements.

%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
%this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
\subsection{Other empirical work}
There is little empirical eassessment of these theories about speculation and noise-trading. In part this is because information is limited and it is usually difficult to distinguish between speculators and other investors and, even more so, between informed and uninformed.

One approach has been to assume that noise-traders utilise simple investment strategies that involve momentum or reversals.  For example, \citet{  shillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation. Alternatively, speculation can be identified with momentum trading.   The literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.  These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.

\subsection{Why this is important}
The traditional Keynesian view of speculation asserted that it would be an additional noise that would make it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. This view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

@@

For example, the debate about a tax on financial transactions, \emph{the Tobin tax} \citet{TobinTax}, has been intended to reduce speculation.  The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.  This argument depends on the idea that speculators will drive the price away from its fundamental value.

@@

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the  foreign exchange market. This will be use to gain a fuller understanding of speculative behaviour.   This will involve three steps:  first, assess the price implication of the noise-trader model;  second, measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.  
% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citep{FrootChartists} for a survey.  %this may go in the explanation for why fx is used.  Use this elsewhere. Add more?

The \citep{Delong1990noise} noise-trader model used in this study. This is a derivative of the \citep{Samuelson1958} overlapping generations model  with two-period agents making a portfolio choice.  There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.  There are two types of investor or trader: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

@@

In words, the deviation from the true fundamental value of unity is a result of the discounted value of the noise-traders' misperception of the price $(p_t - \rho^*)$, multiplied by the weight of noise-traders in the total $(\mu)$ plus a perpetuity of the average noise-trader misperception $(\frac{ \mu \rho^*}{r} )$, adjusted for the risk.   This risk is a function of the volatility of the price which,  from equation \ref{eq:NP1} is a result of the variability of the misperception of speculative noise-traders.   All but the second term in equation \ref{eq:NP1} are constants, and if the variance of the price one period ahead  is equal to the variance of the misperception,

@@

The following effects are noted:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief \citep[pp. 14-15]{Delong1990noise}. 
%Put something on these four in the conclusion. 
The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.  While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.  The noise-traders expected to get higher utility due to their misperception.  The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.  This is consistent with the type of over-confidence that has been found by \citet{odean1998volume}in day-traders.

@@

To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of mis-perceptions and noise-trader weights.  Two time series are created.  The first series is based on a random noise-trade misperception, $(\rho^*)$.  This is called the Misperception Model (MPM). The second is based on a random weight of noise-traders $(\mu)$ and is called the Random Weight Model (RWM).   The simulation creates a 1000 period price series from Equation \ref{eq:NP2}. The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citep[p. 127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader mis-perception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when mis-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study. The R package \emph{eventstudies} createed by \citet{eventstudies} is used for the data manipulation.  Descriptive statistics for the whole series of returns and for those when misperception or weight of noise-traders are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

@@

\end{tabular}
\begin{tablenotes}
\small
\item Descriptive statistics for the De Long Noise-Trader model \citep{Delong1990noise}.  The Base Case measures the benchmark as returns from prices generated by Equation \ref{eq:NP2} for the whole sample.   The MPM measures returns after a period of extreme sentiment when the misperception of noise-traders ($\rho$) is equal to or above the 95th percentile of all the misperceptions; the RMW measures the returns following a recording of extreme weight of noise-traders ($\mu$) is equal to or above the 95th percentile of all noise-trader realisations.   The asterisk denotes significantly different from zero at the 95th confidence interval.    
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction. In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.%\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.  However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.   The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.  The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.  In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons. Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.  Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citep{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
This is the relative price of currency put and call options for the same strike price.  The \emph{Put-Call-Parity} in the Black-Scholes option pricing model assumes that these will be the same \citep{Black1973Options}.  While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citep{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.

@@

The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.  This date is shortly before this part of the research was started. The Euro data runs  from March 3 1998 to December 31 2008.  There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.  The absence of these data seems to be a result of problems with the CFTC database.  The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.  For the CoT data, two series are created:  S1 measures the net long non-commercial positions relative to total non-commercial positions and S2 measures the net long non-commercial positions relative to open interest.  Table \ref{tabref:DS3} given an overview of the descriptive statistics for the returns on the weekly foreign exchange rates that are relevant for the futures market data.  There are no significant differences from the daily data that is used in the risk-reversal part of the study recorded in Table \ref{tabref:DS1}.  Once again, the mean return is very close to zero.  This is consistent with the idea that exchange rates follow a random walk.  The distribution of returns are skewed for some pairs and often fat-tailed (identified with an asterisk where statistically significant as either a t-test or a test of the standard error of skew or standard error of kurtosis as calculated above).

>>>>

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study is used to assess the views of Keynes and Friedman on speculation. Speculative extremes are ranked by intensity of sentiment and weight of speculative activity. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.  The extremes say little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity. 
\end{abstract}
\section*{Introduction}
There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of this spectrum is the view of Keynes from Chapter 12 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Here speculation is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that does not emerge randomly, but feeds off itself in an evolution of collective sentiment that encourages trends in asset prices.

Towards the other end of the spectrum, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
    The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measurd in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing the the strength of speculative opinion or the weight of speulative positions does not improve forecasts of future returns. This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.

\section{Literature}
\subsection*{Overview}
  Speculation is an important concept in economics, particular in the area of finance.  However, there is no clear, unambiguous theory of speculation. Speculation plays a key role in three important and related areas: market-microstructure, the rationale for the trading of financial securities and rational expectations.

@@

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninforrmed speculation provides a rational and an opportunity for the trading of financial securities. If all economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, the informed can take advantage of their information.  This creates a justification for acquiring information. Speculative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundametnal value is thereby essential to encourage te process of price discovery. 
%
 \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

  Speculation is also an important part of the argument for why prices should tend towards a level that consistent with \emph{the rational expectations equilibrium} (RER). This is the level that is consistent with rational expectations.  In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent. These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the means by which the components of the model tend towards equilibrium values.  
% See \citet gc{Lyons2006Microstructure} and \citep{Evans2002Order} for examples of the micrstrucure approach in foreign exchange.

However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inapprpropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements.

The institutional explanations are assciated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liqudiity require larger margins and deeper haircuts; investors liquidate their fund holdings; the atttraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.  Psychological explanations revolve around the common panic, the group-thining and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than acertain underlying, fundamental value \citet[p.101]{Keynes1936}

it has also been suggested that speculative technieques can change overtime.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation. 
%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
%this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
\subsection{Empirical work}
Empirical assessment of speuclation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish those that may be noise-traders from those assumed to be informed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  For example speculation can be identified with momentum trading. It is also possible to assume that certain investors, such as investment banksa and hedge funds, are informed while other trade without inforation. The latter group may include day-traders or private individuals.

%The literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.  These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.

\subsection{Why this is important}
The Keynesian view of speculation is that it is an addition to market noise that makes it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. This description provides an explanation for Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

@@

The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax} , has been intended to reduce speculation.  The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.  The argument for a tax on finanical transactions depends on the idea that speculators will drive the price away from its fundamental value.

@@

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the  foreign exchange market. This will be use to gain a fuller understanding of speculative behaviour.   The experiement will involve three steps:  assess the price implication of the noise-trader model;  measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.  
% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citet gc{FrootChartists} for a survey.  %this may go in the explanation for why fx is used.  Use this elsewhere. Add more?

The \citet{Delong1990noise} noise-trader model used in this study. This is a derivative of the \citet{Samuelson1958} overlapping generations model  with two-period agents making a portfolio choice.  There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.  There are two types of investor or trader: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

@@

The deviation from the true fundamental value of unity is a result of the discounted value of the noise-traders' misperception of the price $(p_t - \rho^*)$, multiplied by the weight of noise-traders in the total $(\mu)$ plus a perpetuity of the average noise-trader misperception $(\frac{ \mu \rho^*}{r} )$, adjusted for the risk.   This risk is a function of the volatility of the price which,  from Equation \ref{eq:NP1} is a consequence of the variability of the misperception of speculative noise-traders.   All but the second term in equation \ref{eq:NP1} are constants, and if the variance of the price one period ahead  is equal to the variance of the misperception,

@@

The following four effects are noted by \citet[pp. 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} as the speculative noise-traders drive prices towards their misperceived vaue; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the noise-trader risk caused by the difference between their perception and the actual value, it is an added risk that is caused by the variability of noise-trader belief as noise-traders could drive the price even further from its fundamental value. In this model, tHe more numerous the noise-traders, the more volatile prices.  
%Put something on these four in the conclusion. 
The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citet{FriedmanPositive}.  While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.  The noise-traders expected to get higher utility due to their misperception.  The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.  This is consistent with the type of over-confidence that has been found by \citet{odean1998volume}in day-traders.

@@

To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of mis-perceptions and noise-trader weights.  Two time series are created.  The first series is based on a random noise-trade misperception, $(\rho^*)$.  This is called the Misperception Model (MPM). The second is based on a random weight of noise-traders $(\mu)$ and is called the Random Weight Model (RWM).   The simulation creates a 1000 period price series from Equation \ref{eq:NP2}. The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citet[p. 127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader mis-perception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.   
% it is possible to have other specifications. For example, the noise-traders oculd become more prominent if their strategy is more successful,

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when mis-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study. The R package \emph{eventstudies} createed by \citet{eventstudies} is used for the data manipulation.  Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

@@

\end{tabular}
\begin{tablenotes}
\small
\item Descriptive statistics for the De Long Noise-Trader model \citet{Delong1990noise}.  The Base Case measures the benchmark as returns from prices generated by Equation \ref{eq:NP2} for the whole sample.   The MPM measures returns after a period of extreme sentiment when the misperception of noise-traders ($\rho$) is equal to or above the 95th percentile of all the misperceptions; the RMW measures the returns following a recording of extreme weight of noise-traders ($\mu$) is equal to or above the 95th percentile of all noise-trader realisations.   The asterisk denotes significantly different from zero at the 95th confidence interval.    
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction. In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.%\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.  However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.   The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.  The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.  In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons. Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.  Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citet gc{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
Risk-reversal skew is the relative price of currency put and call options for the same strike price.  The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same.  While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.

@@

The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.  This date is shortly before this part of the research was started. The Euro data runs  from March 3 1998 to December 31 2008.  There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.  The absence of these data seems to be a result of problems with the CFTC database.  The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.
